---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mochita

<!-- badges: start -->
<!-- badges: end -->

mochita lets you write integration tests for your R web applications.
You declare an HTTP request and the expectations its response should
meet; `$perform()` then spins up your app on a random local port, sends
the request for real, and checks the response with
[testthat](https://testthat.r-lib.org/).

## Installation

You can install the development version of mochita from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("JulioCollazos64/mochita")
```

## Example

`mochita()` takes an app definition as understood by
[httpuv](https://github.com/rstudio/httpuv), i.e. a list with a `call`
function that receives a Rook request and returns the response:

```{r app}
app <- list(
  call = function(req) {
    list(
      status = 200L,
      headers = list("Content-Type" = "text/plain"),
      body = paste0("hello, ", req$HTTP_USER_AGENT)
    )
  }
)
```

Pick an HTTP method, build up the request, declare the expectations and
finally `$perform()` it:

```{r example}
library(mochita)

mochita(app)$get("/")$
  set("User-Agent", "mochita")$
  expect(200)$
  expect("Content-Type", "text/plain")$
  expect("hello, mochita")$
  perform()
```

## Expectations

`$expect()` figures out what to check from the arguments you give it:

``` r
mochita(app)$get("/")$
  set("User-Agent", "mochita")$
  expect(200)$                          # response status
  expect(200L, "hello, mochita")$       # status and body at once
  expect("Content-Type", "text/plain")$ # a header name/value pair
  expect("hello, mochita")$             # the response body
  perform()
```

For anything else, pass a function taking the response — a list with
`status`, `headers` and `body` — and use any testthat expectation
inside:

``` r
mochita(app)$get("/")$
  set("User-Agent", "netscape")$
  expect(function(res) {
    testthat::expect_match(res$body, "netscape")
  })$
  perform()
```

## Sending a request body

Use `$send()` to attach a body to the request, for instance when
testing a JSON endpoint:

```{r json}
json_app <- list(
  call = function(req) {
    input <- yyjsonr::read_json_raw(req$rook.input$read())

    body <- yyjsonr::write_json_str(
      list(x = seq_len(input$len)),
      opts = yyjsonr::opts_write_json(auto_unbox = TRUE)
    )

    list(
      status = 200L,
      headers = list("Content-Type" = "application/json"),
      body = body
    )
  }
)

mochita(json_app)$post("/json")$
  set("Accept", "application/json")$
  send('{"len":3}')$
  expect("Content-Type", "application/json")$
  expect(200L, '{"x":[1,2,3]}')$
  perform()
```

## Reusing the app

Passing the app each time is not necessary. If you are testing the same
app, assign the result of `mochita()` once — each method verb starts a
fresh request:

```{r reuse}
req <- mochita(app)

req$get("/")$set("User-Agent", "firefox")$expect("hello, firefox")$perform()

req$get("/")$set("User-Agent", "chromium")$expect("hello, chromium")$perform()
```

## Use in your test suite

Since expectations run through testthat, mochita calls slot directly
into your `tests/testthat/` files:

``` r
describe("GET /", {
  it("greets the user agent", {
    mochita(app)$get("/")$
      set("User-Agent", "mochita")$
      expect(200)$
      expect("hello, mochita")$
      perform()
  })
})
```
